---
title: "syntactic"
author: "Michael Steinbaugh"
date: "`r BiocStyle::doc_date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{syntactic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

# Introduction

```{r}
library(syntactic)
data(syntactic, package = "acidtest")
```

syntactic improves upon the `make.names()` functionality defined in base R, and is designed to return syntactically valid names from biological metadata. The package exports these primary functions:

- `camelCase()` (e.g. `helloWorld`).
- `dottedCase()` (e.g. `hello.world`).
- `snakeCase()` (e.g. `hello_world`).
- `upperCamelCase()` (e.g. `HelloWorld`).

syntactic is designed to handle many common mixed case acronyms (e.g. mRNA, RNAi), as well as decimals in names.

# String mode

`camelCase`
`dottedCase`
`snakeCase`
`upperCamelCase`

Additionally, the package exports these string utility functions:

- `makeNames()`: modified variant of `make.names()` that sanitizes using underscores instead of dots.
- `capitalize()`: Capitalize the first letter of all words in a character vector.

`capitalize`
`sentenceCase`

`makeNames`
`makeDimnames` works on row and column names for a matrix or data frame.

Plot utilities:

`makeLabel`
`makeTitle`
`makeWords`

# File rename mode

`snakeCase`.
`kebabCase`.

`camelCase` and `upperCamelCase` are also supported, but not generally recommended for file names.


# Related packages

If syntactic doesn't work quite right for your workflow, these popular packages also provide excellent sanitization support:

- [janitor][] by Sam Firke.
- [lettercase][] by Christopher Brown.
- [snakecase][] by Malte Grosser.

# R session information

```{r}
utils::sessionInfo()
```

[basejump]: https://basejump.acidgenomics.com/
[bioconda]: https://bioconda.github.io/
[conda]: https://conda.io/
[janitor]: https://cran.r-project.org/package=janitor
[lettercase]: https://cran.r-project.org/package=lettercase
[r]: https://www.r-project.org/
[snakecase]: https://cran.r-project.org/package=snakecase
[syntactic]: https://syntactic.acidgenomics.com/
